A Large-Scale Study of the Use of Eval in R

### 2020-08-28 inst/paper/main.tex
Add stuff

\subsubsection{Provenance}

In this section we turn our attention to the \AllExpressionInputEvalCallPerc
expressions passed as input to \eval. There are multiple ways to obtain
expressions in R.

\c{eval} takes an R expression. To evaluate a string, it first has to be parse
using either \c{parse}, \c{str2expression} or \c{str2lang} functions. The
difference between them is that the latter two are specialized versions of
\c{parse} that do not keep source references (line and column numbers). This
happens for \PercentParsedEvals of the eval calls, and \PercentParsedCallSites
of the call sites. The input to \c{parse} can come from either a string, a file
or directly a from a user input in interactive REPL session\footnote{in the
  corpus, there is only one call site to eval that takes parse expression from a
  user input, \c{invokeRestartInteractively} that prompts user for new argument
  values when a function invocation fails.}.

This corresponds to strings loaded from a file or to metaprogramming, \ie
building new variables or functions by concatenating strings together, as in
Listing~\ref{code:parsepaste}. To detect this kind of metaprogramming, we look
at the usage of the main concatenating functions in R, \c{paste}, \c{paste0},
\c{glue} and \c{str_c}, and \PasteParsedEvals of the evals and
\PasteParsedCallSite of the call sites using a parse function actually call a
concatenating function.

\begin{lstlisting}[caption={Using \c{eval} and \c{parse} to do meta-programming in package \emph{data.table}, to build the options at load time (in \emph{onLoad.R}).}, label=code:parsepaste]
eval(parse(text=paste0("options(",i,"=",opts[i],")")))
\end{lstlisting}

\textbf{What about a table of packages ranked by usage of parse (and/or paste...)?}

>>>>

We looked at the top ten expressions passed to core and
package \eval calls.
For package \eval calls, the ten most frequently evaluated expressions account
for 80\% of all package eval calls. The most frequently occurring call is to the
\eval in \c{ggplot2::ggproto} which happens over 68\% of the times. The
remaining calls

%%eval(quote(function(value) freduce(value, `_function_list`)),
%%env, env), c("fseq", "function")
\subsubsection{Provenance}

In this section we turn our attention to the \AllExpressionInputEvalCallPerc
expressions passed as input to \eval. There are two ways to obtain
expressions in R.
\begin{itemize}
  \item \c{substitute} function can be used to synthesize ASTs from
  expressions by replacing symbols with their bindings in the specified
  environment.
  \item \c{expression} function can be used to create a vector of expression
    objects from text.
  \item \c{parse} function and its variants \c{str2expression} and \c{str2lang}
    can be used to turn strings or text from file into R expressions.
\end{itemize}

In our corpus, we observe 5.2\% cases where \eval directly evaluates output of
\c{substitute}, 0.7\% cases where output of \c{parse} and its variants is read
directly and only 704 cases for \c{expression}. Majority of expression objects
consumed by \eval are generated by their transitive uses by other functions.

The pattern \c{eval(parse(...))} can be used for dynamic code loading. This
forms the core of \c{source} and \c{sys.source} functions in R that are commonly
used for loading code in R files in interactive settings. We investigated the
number of cases in which the output of \c{parse} and its variants is passed to
\eval, directly or transitively by tainting their output. We observed that
\PercentParsedEvals of the eval calls consume the result of calling \c{parse} on
a file and \PasteParsedEvals of the eval calls consume the result of calling
\c{parse} on a string. This corresponds to \PercentParsedCallSites and
\PasteParsedCallSite of the total \eval call sites. We also identified one
function in core R, \c{invokeRestartInteractively} that prompts the user for
input, parses it, and passes it to \eval.
